---
title: "HOMEWORK - 6 - SMT"
author: "Ümmü Savran"
date: "2024-04-18"
output: pdf_document
---

```{r}
# Load the necessary libraries
library(ggplot2)
library(dplyr)

```

## DATA SET

```{r}
# Load required libraries
library(MASS)

# Generating the dataset
set.seed(123) # Set seed for reproducibility


# Define sample size
n <- 100

# Generate age data: random values between 20 and 60
age <- sample(20:60, n, replace = TRUE)

# Generate experience data: random values between 1 and 30
experience <- sample(1:30, n, replace = TRUE)

# Generate main dish service time data: random values following a normal distribution with more variability
service_time <- rnorm(n, mean = 20 + 0.5*age - 0.2*experience, sd = 5)

# Create the dataset
data <- data.frame(Age = age, Experience = experience, Service_Time = service_time)

# Check the summary of the dataset
summary(data)

# Define age and experience levels
data$Age_Group <- ifelse(data$Age < 40, "Low", "High")
data$Experience_Group <- ifelse(data$Experience < 20, "Low", "High")

```

```{r}
data
```


## REGRESSİON MODEL

```{r}
# Fit a linear regression model
model <- lm(Service_Time ~ ., data = data)

# Summary of the regression model
summary(model)

```

In this model, a linear regression model was applied to evaluate the effects of independent variables (Age, 
Experience, Age_GroupLow, Experience_GroupLow) on Service_Time. When the p-values of the model are examined, 
it is seen that the age variable has a statistically significant effect on Service_Time (p < 0.001). In other
words, it can be said that as age increases, the length of service also increases. However, it cannot be said
that the Experience variable (p = 0.7485) and the age group (Age_GroupLow) and experience group 
(Experience_GroupLow) variables (both p > 0.05) have a statistically significant effect. The effect of these 
variables on Service_Time is not obvious. The R-squared value of the model is 0.5764, which indicates that 
the independent variables explain 57.64% of the total variance. However, the corrected R-squared value is 
0.5586, which may indicate that the model has less explained variance and perhaps the model can be improved 
with more independent variables. The f-statistic is 32.32 and the p-value is < 2.2e-16, which indicates that 
the model is significant, that is, there is statistically significant evidence that at least one argument 
describes Service_Time However, further evaluation may be required to determine exactly how well the model 
fits.

## ANOVA ANALYSIS

```{r}
# Perform ANOVA
anova_result <- anova(model)

# Summary of ANOVA results
print("ANOVA Results:")
print(anova_result)

```

While these ANOVA results show that age and experience variables have a significant effect on Service_Time, 
they show that the effect of age group and experience group is not statistically significant. The effect of 
age and experience variables on Service_Time is quite noticeable, but the effect of age group and experience 
group variables is weak or almost nonexistent. These results show that SERVICE_TIME depends more precisely on
individual characteristics such as age and experience, but age and experience groups do not make a 
significant difference in this context.

## ANCOVA ANALYSIS

```{r}
# Perform ANCOVA
ancova_result <- lm(Service_Time ~ Age + Experience, data = data)

# Summary of ANCOVA results
summary(ancova_result)


```

These ANCOVA results show that age and experience variables have a significant impact on Service_Time. The 
effect of age on Service_Time is quite pronounced (p < 0.001), Service_Time tends to increase as age 
increases. The effect of the experience variable is statistically significant (p = 0.015), but this effect is
weaker than the effect of age. These results show that SERVICE_TIME depends on factors such as both age and 
experience. However, the effect of the age variable is stronger than the experience variable.


```{r}
# Visualize the data using scatterplot
plot(data$Age, data$Service_Time, 
     xlab = "Age", ylab = "Service Time", 
     main = "Scatterplot of Service Time vs Age")

```

This Decoupling graph visualizes the relationship between age and length of service. On the graph, there 
seems to be a trend that as the age increases, the length of service usually increases as well. However, it 
is noticed that there are large fluctuations in the duration of service in some age values. These 
fluctuations may indicate that the relationship between age and length of service is influenced by some other
factors, or they may affect the accuracy of the model. Dec. These points should be examined in more detail.


